Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-02-01 Manuscript/manuscript.Rmd
manuscript edits

In this study we have developed an ensemble framework for predicting TAD boundaries that addresses class imbalance problem. We evaluated Hi-C data for the GM12878 cell line obtained at 10 kb, 25 kb, 50 kb, and 100 kb resolution. For each resolution, we examined the inclusion of three different predictor types, including distance as the measure of association between genomic elements. We then tested four re-sampling techniques aimed at creating balanced classes. Random forest classification was performed to both compare each combination of the above ensemble, and investigate which functional genomic features were most associated with the TAD boundaries.  Model performance was evaluated using accuracy, area under the precision-recall curve (AUPRC), F1 score, and Matthew's Correlation Coefficient (MCC). This process was then repeated for X additional cell lines including K562, ......

@@

The efficacy of different ensemble systems was compared using various performance metrics including accuracy, F1-score, and MCC. These metrics are defined as follows:

$$Accuracy = \frac{TP + TN}{TP + TN + FP + FN}$$
$$F1 - Score = \frac{2TP}{2TP + FP + FN}$$
$$MCC = \frac{TP \times TN - FP \times FN}{\sqrt{ (TP + FP)(TP+FN)(TN+FP)(TN+FN) }}$$

@@

For the task of identifying the functional genomic elements that were most predictive of the formation of TAD boundaries, we evaluated the variable importances associated with each predictor of the random forest models associated with each resampling technique. For each model and within each tree, the prediction accuracy on the out-of-bag portion of the data was recorded. Then the same was done after permuting each predictor variable. The difference between the two accuracies was then averaged over all trees, and normalized by the standard error. All random forest algorithms were implemented in R using the `caret` package version 6.0 [@kuhn2012caret].

>>>>

In this study we have developed an ensemble framework for predicting TAD boundaries that addresses class imbalance problem. We evaluated Hi-C data for both the GM12878 and K562 cell line obtained at 10 kb, 25 kb, 50 kb, and 100 kb resolution. For each resolution, we examined the inclusion of three different predictor types, including distance as the measure of association between genomic elements. We then tested four re-sampling techniques aimed at creating balanced classes. Random forest classification was performed to both compare each combination of the above ensemble, and investigate which functional genomic features were most associated with the TAD boundaries.  Model performance was evaluated using area under the precision-recall curve (AUPRC), F1-Score, and MCC.

@@

The efficacy of different ensemble systems was compared using the F1-score and MCC. These metrics are defined as follows:

$$F1 - Score = \frac{2TP}{2TP + FP + FN}$$
$$MCC = \frac{TP \times TN - FP \times FN}{\sqrt{ (TP + FP)(TP+FN)(TN+FP)(TN+FN) }}$$

@@

For the task of identifying the functional genomic elements that were most predictive of the formation of TAD boundaries, we evaluated the variable importances associated with each predictor of the random forest models associated with each resampling technique. For each model and within each tree, the AUPRC on the out-of-bag portion of the data was recorded. Then the same was done after permuting each predictor variable. The difference between the two AUPRC values was then averaged over all trees, and normalized by the standard error. All random forest algorithms were implemented in R using the `caret` package version 6.0 [@kuhn2012caret].
